---
title: "linear_models"
output: html_document
date: "2025-11-06"
---

```{r}
library(tidyverse)
library(p8105.datasets)

```

Load the NYC airbnb data

```{r}
data(nyc_airbnb)
```

Look at the data / do some cleaning

```{r}

nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename(
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, room_type, neighbourhood)

```

Do regression!!

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

do some additional cleaning then refit

```{r}
nyc_airbnb =
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), #most common = reference group
    room_type = fct_infreq(room_type)
  )

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

look at `lm` stuff

```{r, eval = FALSE}

#clunky and don't recommend doing this
summary(fit)
names(summary(fit))
summary(fit)[["coefficients"]]
summary(fit)[["df"]]

fitted.values(fit)
resid(fit)

```

Look at cleaner `lm` stuff

```{r}
fit |> 
  broom::tidy() |> #pulls out what you care about from linear model
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)

#basically gives you a dataframe/table of the parameter estimates

fit |> 
  broom::glance() #r-squared, adj. r-squared, sigma, test stat, p-value, df, loglik, AIC, BIC
```

## Diagnostics

Look at residuals

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = resid))+
  geom_histogram()

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = borough, y = resid))+
  geom_violin()

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid))+
  geom_point()
```

What about a categorical variable?

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
#can compare these directly head-to-head b/c they are nested

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Interactions vs. nested data

```{r}
fit_interactions =
  lm(price ~ stars*borough + room_type*borough, data = nyc_airbnb)
##don't have to include main effects in R, it gives it to you automatically

fit_interactions |> 
  broom::tidy()
```

Let's just look at Brooklyn first

```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |>  #placeholder dataset goes here "_"
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |>  #placeholder dataset goes here "_"
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |>  #placeholder dataset goes here "_"
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |>  #placeholder dataset goes here "_"
  broom::tidy()

##Uh oh! We copied and pasted code multiple times. Time to create a function!
```

Write a short function

```{r}
#Here's my function:
lm_airbnb = function(df) {
  
  lm(price ~ stars + room_type, data = df)
  
}

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm_airbnb() |> #using my function that only requires one argument
  broom::tidy()
```

Create a list of dataframes, and iterate to fit the model each time

```{r}
nested_lm_results =
  nyc_airbnb |> 
  nest(data = -borough) |>  #data = everything except borough
  mutate(
    fits = map(data, lm_airbnb), #fitting 4! models! all! at! once!
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results)
  
```

Do some untidying

```{r}
nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

Use an *anonymous* function instead of `lm_airbnb`

```{r}
nested_lm_results =
  nyc_airbnb |> 
  nest(data = -borough) |>  
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)), 
    #instead of using lm_airbnb function
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results)
```

Let's do an even more extreme example!!

```{r}
manhattan_analysis =
  nyc_airbnb |> 
  filter(
    borough == "Manhattan"
    ) |> 
  nest(data = -neighbourhood) |> 
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(fits, broom::tidy)
  ) |> 
  select(neighbourhood, results) |> 
  unnest(results)
#fit 32 lm's, one for each neighborhood in Manhattan, look at results in a dataframe
```

Make a plot

```{r}
manhattan_analysis |> 
  filter(term == "stars") |> 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, estimate)
  ) |> 
  ggplot(aes(x = neighbourhood, y = estimate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```





